---
output:
  pdf_document: default
  html_document: default
---
## Sample code for:
## *Last name analysis of mobility, gender imbalance, and nepotism across academic systems*
## by Jacopo Grilli and Stefano Allesina, PNAS 2017

## Data
The data are stored in the `data` directory, which contains a file for each dataset:
```{r}
dir("../data")
```

To read a data file, use either `read.csv` or the much faster `read_csv` from `tidyverse`:
```{r, message = FALSE}
library(tidyverse)
data <- read_csv("../data/ita_2000.csv")
glimpse(data)
```

For each researcher, `last_id` is a numeric code corresponding the researcher's last name. All data sets contain the columns `last_id`, `institution_id`,`city_id`, `region`, and `sector`. The Italian data also contains `gender` and `rank`.

## Randomizations
The code to perform the randomizations illustrated in the article is contained in `run_randomizations.R`. The code requires the packages `data.table` and `tidyverse`. 

To load the code:
```{r, message = FALSE}
source("run_randomizations.R")
```

To launch a randomization, invoke the function `randomize_compute_pval`. This function accepts three parameters:

- `data_file`: the data to use

- `randomization`: the type of randomization to perform. Possible choices are `"nation"`, `"city"`, and `"field"`.

- `nrand`: the number of randomizations to perform (in the article, $10^6$---but for testing use a smaller number, as it might take a long time otherwise)

For example, to randomize the `us_2016.csv` data by shuffling last names within each field, use:
```{r, message = FALSE}
us_2016_byfield <- randomize_compute_pval(data_file = "../data/us_2016.csv", 
                                          randomization = "field", 
                                          nrand = 5000)
```

The object `us_2016_byfield` is a list with containing two tables, `field` and `region`:
```{r}
us_2016_byfield$field
us_2016_byfield$region
```

For each table, `totpairs` is the total number of isonymous pairs observed in the discipline or region, `mean` is the expected number of pairs in the randomized data, `sd` the standard deviation, and `pvalue` the probability of observing a number of pairs in the randomization that is larger or equal than that in the original data.

To visualize the results, simply call 

```{r}
visualize_results(us_2016_byfield)
```

where saturated bars represent significant results once accounted for multiple hypothesis testing.

